---
title: "ggplot"
author: "AMBenson"
date: "October 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import Data and load libraries

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(tidyr)
  library(dplyr)
  library(DT)
})

# https://knb.ecoinformatics.org/#view/urn:uuid:e05865d7-678d-4513-9061-2ab7d979f8e7
# Search 'permit value'
permits <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Aa3c58bd6-481e-4c64-aa93-795df10a4664", method = "libcurl"), stringsAsFactors = FALSE)

datatable(permits)
```

scatter plot 2 dimensional (x and y axis) therefore must know your x and y values
geometrical aspect of the plot (color, size, shape of geom)
aes--aesthetic is bad name!  map variable to geometric aspect of plot
think in terms of mapping--output of the aesthetic mapped to plot
plus sign to add things to plot--layer things
```{r}
help("GeomPoint")
ggplot(permits, aes(x=Value, y=StdDev,color=Gear,shape=Region)) + 
  geom_point()
```
Assigning aesthetics--only for variables
has to be the name of a variable in your data set 
if you assign color= blue in the aesthetic wonkiness--should be in the geom_point
```{r}
ggplot(permits, aes(x=Value, y=StdDev,shape=Region)) + 
  geom_point(color='blue')
```
histogram--only need a single variable
defaults to 30 bins if you don't say something different
If you get a warning, your column may have an NA or missing value
```{r}
ggplot(permits, aes(Value)) + 
  geom_histogram(binwidth = 10000)
```
Boxplot##tells you so much infor about your data
```{r}
ggplot(permits, aes(Gear, Value)) + 
  geom_boxplot()
```

```{r}
permits_ci <- permits %>%
  filter(Region == "Cook Inlet")

ggplot(permits_ci, aes(Gear, Value)) + 
  geom_violin()+
   geom_point(shape = 1, position = "jitter")
```

```{r}
ggplot(permits, aes(Value, StdDev)) + 
  geom_point() +
  geom_smooth()
```
note the grey area plots the 95% confidence
```{r}
ggplot(permits, aes(Value, StdDev)) + 
  geom_point() +
  geom_smooth(method='lm')
##default method is loess
```
Summarize data and the plot a bar plot
position and stat are the fundamental things to understand
position--controls how regions displayed stack=put them on top of eachother
stat--default is mean

```{r}
permits_sum <- permits %>% 
  group_by(Gear, Region) %>% 
  summarize(mean_value = mean(Value, na.rm = T))

ggplot(permits_sum, aes(x = Gear, y = mean_value, fill = Region)) +
  geom_bar(position = "stack", stat = "identity")
```
Line graph
scale_y_continous....change the format of the y axis
```{r}
###filter only SE AK permits
permits_se_seine <- permits %>% 
  filter(Gear == "Purse Seine", 
         Region == "Southeast")

ggplot(permits_se_seine, aes(Year, Value)) + 
  geom_point() +
  geom_line()
  scale_y_continuous(limits = c(0, max(permits_se_seine$Value)))+
  
```
style/color preferences
```{r}
ggplot(permits, aes(Value, StdDev, color = Gear)) + 
  geom_point() + 
  scale_color_hue(h = c(270, 360)) # blue to red
  
```
themes
```{r}
ggplot(permits, aes(Value, StdDev, color = Gear)) + 
  geom_point() +
  theme_classic()
```
```{r}
ggplot(permits, aes(Value, StdDev, color = Gear)) + 
  geom_point() +
  theme_classic() +
  theme(legend.position = "bottom",
        legend.background = element_rect(fill = "#EEEEEE", color = "black"),
        legend.title = element_blank(),
        axis.title = element_text(size = 16))
```

```{r}
ggplot(permits, aes(Value, StdDev, color = Gear)) +
  geom_point() +
  theme_classic() +
  theme(legend.position = c(1, 1),
        legend.justification = c(1,1),
        legend.direction = "horizontal",
        legend.title = element_blank()) +
  xlab("Permit Value (unadjusted USD)") +
  ylab("Permit Std. Dev. (unadjusted USD)") +
  ggtitle("Permit Standard Deviation against Permit Value by Gear Type",
          "or: I wish I owned a seine permit")
```
```{r}
my_theme <- theme_bw() +
  theme(legend.title = element_blank(),
        panel.grid.major = element_line(size = 1, linetype = 4),
        panel.grid.minor = element_blank())
  
ggplot(permits, aes(Value, StdDev, color = Gear)) +
  geom_point() +
  my_theme

```

```{r}

```

facets##repeats of the same plot 
```{r}
ggplot(permits_ci, aes(Year, Value)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ Gear)
```
facet_wrap() chose a layout for us but, in this case, it aids comparison if we stack eachpanel on top of one another:
```{r}
ggplot(permits_ci, aes(Year, Value)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ Gear, ncol = 1)
```

```{r}
suppressPackageStartupMessages({
  library(gridExtra)
})

p1 <- ggplot(permits_se_seine, aes(Year, Value)) +
  geom_point() + 
  geom_line()

p2 <- ggplot(permits %>% filter(Gear == "Purse Seine"), aes(Region, Value)) +
  geom_boxplot() +
  scale_y_continuous(labels = function(x) { format(x, scientific = FALSE) })

grid.arrange(p1, p2)
```

